BSOS

### 2014-06-09 Paper.tex
center code

Unfortunately, RISC is not without its downsides as well. Because RISC instruction sets contain opcodes that are designed to execute within a single memory cycle, they often outpace their own hardware and are bottlenecked by latency when reading from and writing to RAM. Many RISC CPUs compensate for this by providing a large on-die L1 cache for compilers and programmers to optimize with, but RAM access is still inefficient in relative to comparable CISC CPUs. Another significant downside is \textit{code expansion}. When high level languages are compiled down to assembly, their advanced control structures and concepts must be translated into the native structures of the system. For example, a high level procedure like this:
\begin{lstlisting}
int a = 5, b = 3;
swap(a, b);
\end{lstlisting}
might be expanded to something like this:
\begin{lstlisting}
mov ax, 5
mov bx, 3
xor ax, bx
xor bx, ax
xor ax, bx
\end{lstlisting}
The end result is an expansion of two high-level instructions into five low-level opcodes, an increase of 250\%. While this happens on all architectures simply due to the nature of abstraction, it is particularly pronounced on RISC systems due to the lack of complex, comprehensive opcodes.

>>>>

Unfortunately, RISC is not without its downsides as well. Because RISC instruction sets contain opcodes that are designed to execute within a single memory cycle, they often outpace their own hardware and are bottlenecked by latency when reading from and writing to RAM. Many RISC CPUs compensate for this by providing a large on-die L1 cache for compilers and programmers to optimize with, but RAM access is still inefficient in relative to comparable CISC CPUs. Another significant downside is \textit{code expansion}. When high level languages are compiled down to assembly, their advanced control structures and concepts must be translated into the native structures of the system. For example, a high level procedure like this:
\begin{center}
\begin{tabular}{c}
\begin{lstlisting}
int a = 5, b = 3;
swap(a, b);
\end{lstlisting}
\end{tabular}
\end{center}
might be expanded to something like this:
\begin{center}
\begin{tabular}{c}
\begin{lstlisting}
mov ax, 5
mov bx, 3
xor ax, bx
xor bx, ax
xor ax, bx
\end{lstlisting}
\end{tabular}
\end{center}
The end result is an expansion of two high-level instructions into five low-level opcodes, an increase of 250\%. While this happens on all architectures simply due to the nature of abstraction, it is particularly pronounced on RISC systems due to the lack of complex, comprehensive opcodes.
